
clear environment and load packades

```{r, load packages}
rm(list=ls())

source("efmody_functions_folder/load_efmody_libraries_and_sources_function.R")

load_efmody_libraries_and_sources_function()

load_efmody_files_function(
  previous_file_name="file_2",
  parent_directory="./efmody_data_files_to_load/",
  list_of_data_frames_to_load=list(
    "all_patient_spline_fits_both_markers"
    ))

```

we could go here for mean absolute difference as well
could go for relative mean absolute difference - frank didn't mention that
Or GINI's mean difference


create a dplyr summarise function so that I can list columns to summarise, and then apply it to different frames and different groupings
```{r, create function to calculate summary statistics}
dplyr_summarise_my_columns <- function(
    data_frame
){
    dplyr::summarise(.data=data_frame,
#####################################################
#17OHP spline fit statistics collated across patients
#17OHP area under curve                     
    mean_auc_17OHp = mean(area_under_curve_17OHP, na.rm=T),
    median_auc_17OHp = median(area_under_curve_17OHP, na.rm=T),
    sd_auc_17OHP = sd(area_under_curve_17OHP, na.rm=T),
#17OHP natural log area under curve
    mean_ln_area_under_curve_17OHP=mean(ln_area_under_curve_17OHP, na.rm=T),
    median_ln_area_under_curve_17OHP=median(ln_area_under_curve_17OHP, na.rm=T),
    sd_ln_area_under_curve_17OHP=sd(ln_area_under_curve_17OHP, na.rm=T),
#17OHP minimum spline values
    mean_min_non_negative_17OHP_spline_fit=mean(min_non_negative_17OHP_spline_fit,na.rm=T),
    median_min_non_negative_17OHP_spline_fit=median(min_non_negative_17OHP_spline_fit,na.rm=T),
    sd_min_non_negative_17OHP_spline_fit=sd(min_non_negative_17OHP_spline_fit,na.rm=T),
#17OHP maximum spline values
    mean_max_non_negative_17OHP_spline_fit=mean(max_non_negative_17OHP_spline_fit,na.rm=T),
    median_max_non_negative_17OHP_spline_fit=median(max_non_negative_17OHP_spline_fit,na.rm=T),
    sd_max_non_negative_17OHP_spline_fit=sd(max_non_negative_17OHP_spline_fit,na.rm=T),
#17OHP sd of spline values
    mean_sd_non_negative_17OHP_spline_fit=mean(sd_non_negative_17OHP_spline_fit, na.rm=T),
    median_sd_non_negative_17OHP_spline_fit=median(sd_non_negative_17OHP_spline_fit, na.rm=T),
    sd_sd_non_negative_17OHP_spline_fit=sd(sd_non_negative_17OHP_spline_fit, na.rm=T),
#17OHP minimum natural log of spline values
    mean_min_non_negative_ln_17OHP_spline_fit=mean(min_non_negative_ln_17OHP_spline_fit,na.rm=T),
    median_min_non_negative_ln_17OHP_spline_fit=median(min_non_negative_ln_17OHP_spline_fit,na.rm=T),
    sd_min_non_negative_ln_17OHP_spline_fit=sd(min_non_negative_ln_17OHP_spline_fit,na.rm=T),
#17OHP maximum natural log of spline values    
    mean_max_non_negative_ln_17OHP_spline_fit=mean(max_non_negative_ln_17OHP_spline_fit,na.rm=T),
    median_max_non_negative_ln_17OHP_spline_fit=median(max_non_negative_ln_17OHP_spline_fit,na.rm=T),
    sd_max_non_negative_ln_17OHP_spline_fit=sd(max_non_negative_ln_17OHP_spline_fit,na.rm=T),
#17OHP sd of natural log of spline values
    mean_sd_non_negative_ln_17OHP_spline_fit=mean(sd_non_negative_ln_17OHP_spline_fit, na.rm=T),
    median_sd_non_negative_ln_17OHP_spline_fit=median(sd_non_negative_ln_17OHP_spline_fit, na.rm=T),
    sd_sd_non_negative_ln_17OHP_spline_fit=sd(sd_non_negative_ln_17OHP_spline_fit, na.rm=T),
#17OHP r squared values, i.e. the coefficient of determination
    mean_ceof_determination_17OHP = mean(coefficient_of_determination_17OHP, na.rm=T),
    median_ceof_determination_17OHP = median(coefficient_of_determination_17OHP, na.rm=T),
    sd_ceof_determination_17OHP = sd(coefficient_of_determination_17OHP, na.rm=T),
#################################
#Androstenedione spline fit statistics collated across patients
#Androstenedione area under curve   
    mean_auc_Androstenedione = mean(area_under_curve_Androstenedione, na.rm=T),
    median_auc_Androstenedione = median(area_under_curve_Androstenedione, na.rm=T),
    sd_auc_Androstenedione = sd(area_under_curve_Androstenedione, na.rm=T),
#Androstenedione natural log area under curve   
    mean_ln_area_under_curve_Androstenedione=mean(ln_area_under_curve_Androstenedione, na.rm=T),
    median_ln_area_under_curve_Androstenedione=median(ln_area_under_curve_Androstenedione, na.rm=T),
    sd_ln_area_under_curve_Androstenedione=sd(ln_area_under_curve_Androstenedione, na.rm=T),
#Androstenedione minimum spline fit values
    mean_min_non_negative_Androstenedione_spline_fit=mean(min_non_negative_Androstenedione_spline_fit,na.rm=T),
    median_min_non_negative_Androstenedione_spline_fit=median(min_non_negative_Androstenedione_spline_fit,na.rm=T),
    sd_min_non_negative_Androstenedione_spline_fit=sd(min_non_negative_Androstenedione_spline_fit,na.rm=T),
#Androstenedeione maximum spline fit values    
    mean_max_non_negative_Androstenedione_spline_fit=mean(max_non_negative_Androstenedione_spline_fit,na.rm=T),
    median_max_non_negative_Androstenedione_spline_fit=median(max_non_negative_Androstenedione_spline_fit,na.rm=T),
    sd_max_non_negative_Androstenedione_spline_fit=sd(max_non_negative_Androstenedione_spline_fit,na.rm=T),
#Androstenedeione SD of spline fit values    
    mean_sd_non_negative_Androstenedione_spline_fit=mean(sd_non_negative_Androstenedione_spline_fit, na.rm=T),
    median_sd_non_negative_Androstenedione_spline_fit=median(sd_non_negative_Androstenedione_spline_fit, na.rm=T),
    sd_sd_non_negative_Androstenedione_spline_fit=sd(sd_non_negative_Androstenedione_spline_fit, na.rm=T),
#Androstenedione minimum spline fit values
    mean_min_non_negative_ln_Androstenedione_spline_fit=
      mean(min_non_negative_ln_Androstenedione_spline_fit,na.rm=T),
    median_min_non_negative_ln_Androstenedione_spline_fit=
      median(min_non_negative_ln_Androstenedione_spline_fit,na.rm=T),
    sd_min_non_negative_ln_Androstenedione_spline_fit=
      sd(min_non_negative_ln_Androstenedione_spline_fit,na.rm=T),
#Androstenedione maximum spline fit values
    mean_max_non_negative_ln_Androstenedione_spline_fit=
      mean(max_non_negative_ln_Androstenedione_spline_fit,na.rm=T),
    median_max_non_negative_ln_Androstenedione_spline_fit=
      median(max_non_negative_ln_Androstenedione_spline_fit,na.rm=T),
    sd_max_non_negative_ln_Androstenedione_spline_fit=
      sd(max_non_negative_ln_Androstenedione_spline_fit,na.rm=T),
#Androstenedione SD of log of spline fit values
    mean_sd_non_negative_ln_Androstenedione_spline_fit=
      mean(sd_non_negative_ln_Androstenedione_spline_fit, na.rm=T),
    median_sd_non_negative_ln_Androstenedione_spline_fit=
      median(sd_non_negative_ln_Androstenedione_spline_fit, na.rm=T),
    sd_sd_non_negative_ln_Androstenedione_spline_fit=
      sd(sd_non_negative_ln_Androstenedione_spline_fit, na.rm=T),
#R squared for androstenedione fit
    mean_ceof_determination_Androstenedione = mean(coefficient_of_determination_Androstenedione, na.rm=T),
    median_ceof_determination_Androstenedione = median(coefficient_of_determination_Androstenedione, na.rm=T),
    sd_ceof_determination_Androstenedione = sd(coefficient_of_determination_Androstenedione, na.rm=T),
####################################
#Comparison of spline fit statistics
#then we collate averages across patients of the comparison between marker spline fits
    mean_mean_ratio_absolute_17OHP_to_Androstenedione = mean(mean_ratio_absolute_17OHP_to_Androstenedione, na.rm=T),
    median_mean_ratio_absolute_17OHP_to_Androstenedione = mean(mean_ratio_absolute_17OHP_to_Androstenedione, na.rm=T),
    sd_mean_ratio_absolute_17OHP_to_Androstenedione = mean(mean_ratio_absolute_17OHP_to_Androstenedione, na.rm=T),
#then we collate averages across patients of the comparison between marker spline fits
    mean_sd_ratio_absolute_17OHP_to_Androstenedione = mean(sd_ratio_absolute_17OHP_to_Androstenedione, na.rm=T),
    median_sd_ratio_absolute_17OHP_to_Androstenedione = median(sd_ratio_absolute_17OHP_to_Androstenedione, na.rm=T),
    sd_sd_ratio_absolute_17OHP_to_Androstenedione = sd(sd_ratio_absolute_17OHP_to_Androstenedione, na.rm=T),
####################################
#Peak and trough statistics
#all patient total peaks 17OHp stats    
    mean_total_peaks_17OHP = mean(total_number_of_peaks_17OHP, na.rm=T),
    median_total_peaks_17OHP = median(total_number_of_peaks_17OHP, na.rm=T),
    sd_total_peaks_17OHP = sd(total_number_of_peaks_17OHP, na.rm=T),
#all patient total trough 17OHP stats
    mean_total_troughs_17OHP = mean(total_number_of_peaks_17OHP, na.rm=T),
    median_total_troughs_17OHP = median(total_number_of_peaks_17OHP, na.rm=T),
    sd_total_troughs_17OHP = sd(total_number_of_peaks_17OHP, na.rm=T),
#all patient total peaks Androstenedione stats
    mean_total_peaks_Androstenedione = mean(total_number_of_peaks_Androstenedione, na.rm=T),
    median_total_peaks_Androstenedione = median(total_number_of_peaks_Androstenedione, na.rm=T),
    sd_total_peaks_Androstenedione = sd(total_number_of_peaks_Androstenedione, na.rm=T),
#all patient total troughs Androstenedione stats
    mean_total_troughs_Androstenedione = mean(total_number_of_peaks_Androstenedione, na.rm=T),
    median_total_troughs_Androstenedione = median(total_number_of_peaks_Androstenedione, na.rm=T),
    sd_total_troughs_Androstenedione = sd(total_number_of_peaks_Androstenedione, na.rm=T),
#summary stats of the different peak and trough values and times across patients:
    mean_value_at_first_peak_17OHP = mean(value_at_first_peak_17OHP, na.rm=T),
    sd_value_at_first_peak_17OHP = sd(value_at_first_peak_17OHP, na.rm=T),
    median_value_at_first_peak_17OHP = median(value_at_first_peak_17OHP, na.rm=T),
    mean_time_at_first_peak_17OHP = mean(time_at_first_peak_17OHP, na.rm=T),
    sd_time_at_first_peak_17OHP = sd(time_at_first_peak_17OHP, na.rm=T),
    median_time_at_first_peak_17OHP = median(time_at_first_peak_17OHP, na.rm=T),
    mean_value_at_first_peak_Androstenedione = mean(value_at_first_peak_Androstenedione, na.rm=T),
    sd_value_at_first_peak_Androstenedione = sd(value_at_first_peak_Androstenedione, na.rm=T),
    median_value_at_first_peak_Androstenedione = median(value_at_first_peak_Androstenedione, na.rm=T),
    mean_time_at_first_peak_Androstenedione = mean(time_at_first_peak_Androstenedione, na.rm=T),
    sd_time_at_first_peak_Androstenedione = sd(time_at_first_peak_Androstenedione, na.rm=T),
    median_time_at_first_peak_Androstenedione = median(time_at_first_peak_Androstenedione, na.rm=T),
    mean_time_at_second_peak_17OHP = mean(time_at_second_peak_17OHP, na.rm=T),
    sd_time_at_second_peak_17OHP = sd(time_at_second_peak_17OHP, na.rm=T),
    median_time_at_second_peak_17OHP = median(time_at_second_peak_17OHP, na.rm=T),
    mean_time_at_second_peak_Androstenedione = mean(time_at_second_peak_Androstenedione, na.rm=T),
    sd_time_at_second_peak_Androstenedione = sd(time_at_second_peak_Androstenedione, na.rm=T),
    median_time_at_second_peak_Androstenedione = median(time_at_second_peak_Androstenedione, na.rm=T)
  )

}

```


```{r, collate summary statistics by arm and degrees of freedom}
spline_summary_stats_total_by_degrees <-
  all_patient_spline_fits_both_markers %>%
  filter(arm!="Screen Failure") %>%
  group_by(degrees_of_freedom, visit_number) %>%
  dplyr_summarise_my_columns()
#add an arm saying 'total' on the left of that frame by creating a simple frame, and binding to it
arm <- data.frame(arm="total")
spline_summary_stats_total_by_degrees <- 
  cbind(arm, spline_summary_stats_total_by_degrees)
#then summarise by other arm
spline_summary_stats_by_arm_and_degrees <-
  all_patient_spline_fits_both_markers %>%
  filter(arm!="Screen Failure") %>%
  group_by(arm, degrees_of_freedom, visit_number) %>%
  dplyr_summarise_my_columns()
#we can then stack the total and the different arms together by binding rows:
spline_summary_stats_by_degrees <-
  rbind(
    spline_summary_stats_total_by_degrees,
    spline_summary_stats_by_arm_and_degrees
  )
```

spline_summary_stats_by_degrees therefore has all the stats we want, just with all the different degrees of freedom, which isn't easy to look down. So we want to filter out each degree of freedom

```{r, print transposed summary stats to csv by each degree of freedom}
#find out how many degrees_of_freedom we have created
total_degrees_of_freedom <- 
  unique(spline_summary_stats_by_degrees$degrees_of_freedom)
dir.create("spline_summary_stats_csv_outputs")
#for each degree of freedom, we want to filter the rows out, then transpose the table and print it to csv

for (each_degree in total_degrees_of_freedom){
  visit_stats_bound_together <- data.frame(NULL)
for (each_visit in 1:4){
  summary_stats_of_degree <- 
    subset(spline_summary_stats_by_degrees, 
           degrees_of_freedom==each_degree & 
           visit_number==each_visit)
  visit_stats_bound_together <- rbind(visit_stats_bound_together, summary_stats_of_degree)
  transposed_summary_stats_of_degree <- as.data.frame(t(summary_stats_of_degree))
  write.csv(transposed_summary_stats_of_degree, 
            paste0("./spline_summary_stats_csv_outputs/transposed_summary_stats_of_degree_", 
                   each_degree, 
                   "_at_visit_number_", 
                   each_visit,
                   ".csv"),
                    row.names=T)
}
  transposed_visit_stats_bound_together <- as.data.frame(t(visit_stats_bound_together))
  write.csv(transposed_visit_stats_bound_together, 
            paste0("./spline_summary_stats_csv_outputs/visit_stats_bound_together_with_", 
                   each_degree,
                   "_degrees_of_freedom",
                   ".csv"),
                   row.names=T)
}
```


that gives us nice tables of the summary stats, but now we want a visualisation of those stats


```{r, create comparison violin plots between arms for a list of metrics}
#find out how many degrees_of_freedom we have created
total_degrees_of_freedom <- 
  unique(spline_summary_stats_by_degrees$degrees_of_freedom)
for (degrees_of_freedom_to_plot in 10){
#filter out the appropriate degrees_of_freedom
spline_summary_data_to_plot <-
  all_patient_spline_fits_both_markers %>%
  filter(arm!="Screen Failure") %>%
  filter(degrees_of_freedom==degrees_of_freedom_to_plot)

parameter_columns_to_plot <- 
  c("area_under_curve_17OHP", 
    "ln_area_under_curve_17OHP",
    "area_under_curve_Androstenedione",
    "ln_area_under_curve_Androstenedione",
    "area_under_curve_absolute_difference",
    "area_under_curve_scaled_difference",
    "max_non_negative_17OHP_spline_fit",
    "max_non_negative_Androstenedione_spline_fit",
    "max_non_negative_ln_17OHP_spline_fit",
    "max_non_negative_ln_Androstenedione_spline_fit",
    "sd_non_negative_17OHP_spline_fit",
    "sd_non_negative_ln_17OHP_spline_fit",
    "sd_non_negative_Androstenedione_spline_fit",
    "sd_non_negative_ln_Androstenedione_spline_fit",
    "sd_ratio_absolute_17OHP_to_Androstenedione",
    "mean_ratio_absolute_17OHP_to_Androstenedione",
    "median_ratio_absolute_17OHP_to_Androstenedione",
    "max_17OHp_to_max_Androstenedione_Ratio",
    "max_17OHP_spline_to_max_Androstenedione_spline_ratio"
    )
#then make the frame long to facilitate box plotting
long_spline_summary_data_to_plot <- 
  pivot_longer(
    spline_summary_data_to_plot, 
    cols = parameter_columns_to_plot,
    names_to = "parameter_to_plot", 
    values_to = "value_to_plot"
  )
#add 'visit number' to the visit number so it facets nicely
long_spline_summary_data_to_plot$visit_name <-
  paste0("Visit Number ", long_spline_summary_data_to_plot$visit_number)
for (parameter in (parameter_columns_to_plot)){
for (visit in 1:4){  
#create a boxplot of what we want
data_to_plot <- subset(long_spline_summary_data_to_plot, 
                     parameter_to_plot==parameter)  
plot_to_print <-
  ggplot(data=data_to_plot, 
               aes(x=arm,
                   y=value_to_plot)) +
  geom_violin(linewidth=0.1,
              draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_signif(comparisons = list(c("Chronocort", "Standard GC therapy")), 
              test="t.test",
              map_signif_level=F,
              vjust=2) +
#  geom_point(size=0.1) +
  geom_dotplot(aes(x=arm,
                   y=value_to_plot),
               stackdir = "center",
               binaxis = "y",
               method = "histodot",
               binwidth = 0.02 * max(data_to_plot$value_to_plot),
               stackratio = 1,
               fill="white",
               dotsize=0.1) +
  facet_wrap(~visit_name) +
#  coord_flip() +

  labs(title=paste0(parameter, "_df_", degrees_of_freedom_to_plot),
       subtitle="Violin lines at quantiles of 0.25, 0.5, 0.75" ,
       y=parameter) + 
  rotate_x_text(60) +
  themepowerpointtitle
dir.create("violin_comparison_plots_by_arm")
dir.create(paste0("violin_comparison_plots_by_arm/degrees_of_freedom_", degrees_of_freedom_to_plot))
ggsave(filename=paste0("violin_comparison_of_", parameter, "_df_", degrees_of_freedom_to_plot, ".tif"), 
       path=paste0("violin_comparison_plots_by_arm/degrees_of_freedom_", degrees_of_freedom_to_plot),
       plot = plot_to_print, 
       device="tiff",  
       width=5, height=10, 
       compression = "lzw", limitsize=F)
  

print("finished rendering parameter =")
print(parameter)
print("finished rendering visit =")
print(visit)
}}}
```

```{r, end of file so save all the listed dataframes into the parent directory}
save_efmody_files_function(
  parent_directory="./efmody_data_files_to_load/",
  parent_file="file_3")
```




